---
title: Assignment A01- GGPLOT2
author: ''
date: '2021-05-21'
slug: dataset-visualisation
categories:
  - ggplot2
tags: []
---

## i) Speed vs Distance
Speed vs Distance relationship through a scatter plot

i)  Speed
ii) Distance

a) The Cars dataset contains the data of cars which covered the distance in mentioned Speed. 
b) The pattern is scattered and clustered in specific linear trend.

```{r warning=FALSE,echo=FALSE,message=FALSE}
library(datasets)
library(ggplot2)
ggplot(cars,aes(x=dist,y=speed)) + geom_point() + labs(title="Speed vs Distance of Cars Dataset",x="Speed (kmph)",y="Distance (km)")
```

## ii) Temperature vs Pressure
Temperature vs Pressure relationship through a scatter plot

i)  Temperature
ii) Pressure

a) The Pressure dataset contains the data of Pressure values measured at the various temperature.
b) The pressure pattern is flat for the initial values of temperature till 200 degree Celsius and increases exponentially after that. 

```{r warning=FALSE,echo=FALSE,message=FALSE}
library(datasets)
library(ggplot2)
ggplot(pressure,aes(x=temperature ,y=pressure)) + geom_point() + labs(title="Temperature vs Pressure of Pressure Dataset",x="Temperature (Celsius)",y="Pressure (Pascal)")
```

## iii) Life Cycle Savings in various countries
DPI vs Count and Pop15 vs Count relationships through a histogram

i)  DPI
ii) Count

a) The LifeCycleSavings dataset contains the data of Life Cycle Savings of various countries.
b) The Savings pattern is shown in the histogram below, which clearly indicates  that the most of the savings are from the DPI lesser than 1000. 

```{r warning=FALSE,echo=TRUE,message=TRUE}
library(datasets)
library(ggplot2)
ggplot(LifeCycleSavings, aes(x = dpi)) + geom_histogram(binwidth = 200) + labs(title="DPI vs Count of LifeCycleSavings Dataset",x="DPI",y="Count")
```

## iv) Eruption in Old Faithful Geyser in Yellowstone National Park
Duration of eruptions vs Waiting time between eruptions relationships through a scatter plot

i)  Duration of eruptions
ii) Waiting time between eruptions

a) The faithful dataset contains the data of Duration of eruptions and Waiting time between eruptions in Old Faithful Geyser in Yellowstone National Park, Wyoming, USA 
b) The timing pattern is shown in the histogram below, which clearly indicates  that the most of the eruption duration falls in the range of 3 to 5. 

```{r warning=FALSE,echo=TRUE,message=TRUE}
library(datasets)
library(ggplot2)
ggplot(faithful, aes(x = eruptions, y = waiting)) + geom_point() + labs(title="Eruptions vs Duration of faithful Dataset",x="Eruptions",y="Duration")
```

## v) Insect Sprays and its count

Spray vs Count relationship through a scatter plot

i)  Spray
ii) Count

a) The InsectSprays dataset contains the data of Insect Sprays and its count.
b) The Spray is concentrated in some some specific count as shown in the ridge plot, which clearly indicates  that the 3 of the density ridges has the count lesser than 10, whereas another 3 of the density ridges has the count distributed in the flat manner from 3 to 30. 

```{r warning=FALSE,echo=FALSE,message=FALSE}
library(datasets)
library(ggplot2)
library(ggridges)
ggplot(InsectSprays, aes(x = count, y = spray, fill = spray, color = spray))+ geom_density_ridges(alpha = 0.5)
```

## vi) Flower Species and its Sepal Length

Flower Species vs Sepal Length relationship through a scatter plot

i)  Species
ii) Sepal Length

a) The iris dataset contains the data of Flower Species and their Sepal Length.
b) The Flowers with various Sepal lengths are categorized under Species such as Setosa, Versicolor and Virginica, which are illustrated by Bar graph. It is clearly seen that the Sentosa has small sepals, Versicolor has bigger sepals than Sentosa and Virginica has the longest sepals in the Iris dataset.

```{r warning=FALSE,echo=FALSE,message=FALSE}
library(datasets)
library(ggplot2)
ggplot(iris, aes(x = Sepal.Length, fill = Species)) +  geom_bar()
```