---
title: Dataset Visualisation
author: ''
date: '2021-05-21'
slug: dataset-visualisation
categories:
  - ggplot2
tags: []
---

## Dataset Visualization Assignment

```{r warning=FALSE,echo=FALSE,message=FALSE}
library(datasets)
datasets::LifeCycleSavings
```

## Life Cycle Savings in various countries
DPI vs Count relationship through a scatter plot

i)  DPI
ii) Count

a) The LifeCycleSavings dataset contains the data of Life Cycle Savings of various countries.
b) The Savings pattern is shown in the histogram below, which clearly indicates  that the most of the savings are from the DPI lesser than 1000. 

```{r warning=FALSE,echo=TRUE,message=TRUE}
library(ggplot2)
ggplot(LifeCycleSavings, aes(x = dpi)) + geom_histogram(binwidth = 300) + labs(title="DPI vs Count of LifeCycleSavings Dataset",x="DPI",y="Count")
```


```{r warning=FALSE,echo=TRUE,message=TRUE}
library(datasets)
datasets::InsectSprays
```
## Insect Sprays and its count
Spray vs Count relationship through a scatter plot

i)  Spray
ii) Count

a) The InsectSprays dataset contains the data of Insect Sprays and its count.
b) The Spray is concentrated in some some specific count as shown in the ridge plot, which clearly indicates  that the 3 of the density ridges has the count lesser than 10, whereas another 3 of the density ridges has the count distributed in the flat manner from 3 to 30. 

```{r warning=FALSE,echo=FALSE,message=FALSE}
library(ggplot2)
library(ggridges)
ggplot(InsectSprays, aes(x = count, y = spray, fill = spray, color = spray))+ geom_density_ridges(alpha = 0.5)
```