---
title: P01 Final Project
author: Raghul Mohankumar
date: '2021-06-22'
slug: p01-final-project
categories:
  - ggplot2
  - Tidyverse
  - shiny
tags: []
---

## Import the Crimes_2020.csv file

```{r warning=FALSE,echo=FALSE,message=FALSE}
library(tidyverse)
library(ggplot2)
library(readr)
crimes <- read_csv("Crimes_2020.csv", na = c(".", ".."))
```

## Description of the Dataset 

```{r warning=FALSE,echo=FALSE,message=FALSE}
cat("The dimension of the dataset is shown as number rows and columns respectively below",dim(crimes))
cat("The columns of the dataset are listed below",names(crimes))
cat("The glimpse of the dataset is shown below")
glimpse(crimes)
cat("The length of the dataset is =",length(crimes))
cat("The compact display of the dataset structure is shown below:")
str(crimes)
cat("The memory size of the dataset is ",object.size(crimes)/1048576, "MegaBytes")
```

## Understanding the Dataset with Transformation functions

```{r warning=FALSE,echo=FALSE,message=FALSE}
crimes_arrange_PT <- crimes %>% arrange(Date, Primary_Type, Case_Number)
crimes_select <- crimes_arrange_PT %>% select(ID:Description)
crimes_filter <- crimes_select %>% filter(Primary_Type == 'THEFT')
cat("The total number of theft cases in the dataset is ","shown below")
count(crimes_filter)
crimes_group_summarize <- crimes %>% group_by(Primary_Type) %>% summarise(n=n()) %>% arrange(desc(n)) %>% top_n(10)
crimes_group_summarize_location <- crimes %>% group_by(Location_Description) %>% summarise(n=n()) %>% arrange(desc(n)) %>% top_n(10)
```

## Understanding the Dataset with Visualization functions


The following visualization depicts that Battery related crimes are highly reported in the year 2020 with the count of 41000 in the city, as per the dataset.Furthermore, the weapon handling crimes are highly reported in 2020, after the battery crimes,  with the count of 40500. However, Robbery is least reported crime in the city.

```{r warning=FALSE,echo=FALSE,message=FALSE}
ggplot(data=crimes_group_summarize, aes(x=Primary_Type,y=n,fill=Primary_Type)) +
  geom_bar(position="dodge",stat="identity", width = 0.5) + 
  ggtitle("Major types of crimes in the dataset ") + scale_x_discrete(guide = guide_axis(n.dodge=3))

```

The following visualization depicts that most of the crimes in the year 2020, are witnessed in the streets with the count of 50000, as per the dataset.Furthermore, the next common location for crimes in 2020 is individual residences with the count of 39000. However, Restaurants are the places with lease crimes with the value of 3000 in 2020, which is mainly due to the preence of surveillance cameras in these locations.

```{r warning=FALSE,echo=FALSE,message=FALSE}

ggplot(data=crimes_group_summarize_location, aes(x=Location_Description,y=n,fill=Location_Description)) +
  geom_bar(position="dodge",stat="identity", width = 0.5) + 
  ggtitle("Major locations of crimes in the dataset") + scale_x_discrete(guide = guide_axis(n.dodge=5))
```
